Temporal ordering and registration of images in studies of developmental dynamics

### 2014-05-28 image_analysis_paper/pnas_draft2.tex
small changes

@@

We presented a unified approach to temporal ordering and registration of images in cross-sectional studies of developmental dynamics. 
%
To the best of our knowledge, algorithmic approaches to these two tasks have primarily been explored independently of each other. 
%
In particular, temporal ordering of imaging data sets was done with a significant amount of human supervision and using registered images as a starting point, e.g., \cite{yuan2014automated, surkova2008characterization}.  
%
In parallel, ordering of large-scale cross-sectional data was done in the context of molecular profiling studies that lacked spatial resolution, in which data are vectors describing the amounts of different chemical species such as mRNA abundances \cite{trapnell2014dynamics, anavy2014blind, gupta2008extracting}. 
%
Automated temporal ordering of such data sets was accomplished by first projecting the data on a low-dimensional subspace spanned by the leading principal or independent components, and then solving a traveling salesman problem or constructing a minimum spanning tree on the projected data to order multiple snapshots. 
%
The approach presented here is different because it addresses image registration and ordering in a single step and orders the data without first projecting them on a low-dimensional subspace. 
%
The task of image registration has also been widely studied \cite{zitova2003image}, for applications such as face recognition \cite{rowley1998rotation}, medical image registration \cite{hajnal2010medical}, and texture classification \cite{greenspan1994rotation}.
%
In contrast to most of the existing approaches to image registration which rely on the knowledge about some landmarks in the data \cite{ian1998statistical} (such as the eyes in face recognition applications \cite{zhao2003face}), algorithms based on angular synchronization (such as vector diffusion maps) can register images even in the absence of such information. 
%TODO: look at Eisen Cell paper about registration

@@

\textbf{Fly strain and whole-mount immunostaining.}
%
Oregon-R (OreR) and spaghetti squash (Sqh)-GFP were used as wild type strains. 
%
{\it Sna\textsuperscript{\it IIG05}/Cyo hb-lacZ} flies were used to study embryos with one peak of dpERK expression.  
%
Embryos were collected and fixed at 22$^\circ$C. 
%
Monoclonal rabbit anti-dpERK (1:100, Cell signaling), mouse anti-Dorsal (1:100, Developmental Studies Hybridoma Bank), rat anti-Twist (1:500, a gift from Eric Wieschaus), and mouse anti-beta-galactosidase (1:500, DSHB) were used to stain proteins of interest. DAPI (1:10,000, Vector Laboratories) was used to visualize nuclei, and Alexa Fluors (1:500, Invitrogen) were used as secondary antibodies.

\textbf{Microscopy.}
%
Nikon A1-RS scanning confocal microscope, and the Nikon 60X Plan-Apo oil objective was used to image embryos. Embryos were collected, stained, and imaged together under the same microscope setting. End-on imaging was performed by using the microfluidics device described previously. Images were collected at the focal plane $\sim$90~$\mathrm{\mu m}$ from the posterior pole of an embryo.

\textbf{Development time during cellularization.} 
%
The progression of membrane invagination during cellularization was used to assign time point to each embryo in the third hour of development. Time-lapse imaging of Sqh-GFP embryo was used to generate correlation curve between the membrane ingression length and the developmental time (n=7) (Figure S1B). The rate at which membrane invaginates was highly reproducible from embryo to embryo. In the experiments with fixed embryos, we measured the membrane length by co-staining embryos with a membrane marker (Sqh-GFP) or by taking phase-contrast images (Figure S1A). We used the correlation curve to assign each embryo to a time window of 2--3~minutes.

>>>>

\significancetext{blahhhhh}

@@

We presented a unified approach to temporal ordering and registration of images in cross-sectional studies of developmental dynamics. 
%
To the best of our knowledge, algorithmic approaches to these two tasks have primarily been explored independently of each other. 
%
In particular, temporal ordering of imaging data sets was done with a significant amount of human supervision and using registered images as a starting point \cite{yuan2014automated, surkova2008characterization}.  
%
In parallel, ordering of large-scale cross-sectional data was done in the context of molecular profiling studies that lacked spatial resolution, in which data are vectors describing the amounts of different chemical species such as mRNA abundances \cite{trapnell2014dynamics, anavy2014blind, gupta2008extracting}. 
%
Automated temporal ordering of such data sets was accomplished by first projecting the data on a low-dimensional subspace spanned by the leading principal or independent components, and then solving a traveling salesman problem or constructing a minimum spanning tree on the projected data to order multiple snapshots. 
%
The approach presented here is different because it addresses image registration and ordering in a single step and orders the data without first projecting them on a low-dimensional subspace. 
%
The task of image registration has also been widely studied \cite{zitova2003image}, for applications such as face recognition \cite{rowley1998rotation}, medical image registration \cite{hajnal2010medical}, and texture classification \cite{greenspan1994rotation}.
%
In contrast to most of the existing approaches to image registration which rely on the knowledge about some landmarks in the data \cite{ian1998statistical} (such as the eyes in face recognition applications \cite{zhao2003face}), algorithms based on angular synchronization (such as vector diffusion maps) can register images even in the absence of such information. 
%TODO: look at Eisen Cell paper about registration

@@

\section{Fly strain and whole-mount immunostaining}
%
Oregon-R (OreR) and spaghetti squash (Sqh)-GFP were used as wild type strains. 
%
{\it Sna\textsuperscript{\it IIG05}/Cyo hb-lacZ} flies were used to study embryos with one peak of dpERK expression.  
%
Embryos were collected and fixed at 22$^\circ$C. 
%
Monoclonal rabbit anti-dpERK (1:100, Cell signaling), mouse anti-Dorsal (1:100, Developmental Studies Hybridoma Bank), rat anti-Twist (1:500, a gift from Eric Wieschaus), and mouse anti-beta-galactosidase (1:500, DSHB) were used to stain proteins of interest. DAPI (1:10,000, Vector Laboratories) was used to visualize nuclei, and Alexa Fluors (1:500, Invitrogen) were used as secondary antibodies.

\section{Microscopy}
%
Nikon A1-RS scanning confocal microscope, and the Nikon 60X Plan-Apo oil objective was used to image embryos. Embryos were collected, stained, and imaged together under the same microscope setting. End-on imaging was performed by using the microfluidics device described previously. Images were collected at the focal plane $\sim$90~$\mathrm{\mu m}$ from the posterior pole of an embryo.

\section{Development time during cellularization} 
%
The progression of membrane invagination during cellularization was used to assign time point to each embryo in the third hour of development. Time-lapse imaging of Sqh-GFP embryo was used to generate correlation curve between the membrane ingression length and the developmental time (n=7) (Figure S1B). The rate at which membrane invaginates was highly reproducible from embryo to embryo. In the experiments with fixed embryos, we measured the membrane length by co-staining embryos with a membrane marker (Sqh-GFP) or by taking phase-contrast images (Figure S1A). We used the correlation curve to assign each embryo to a time window of 2--3~minutes.
